A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-03-01 sections/Discussion.tex
slight modifications to S5

One may blame the inference rule for lambda functions for not collecting garbage correctly.
Here we also show another rule for application inference with a similar problem:
$$
\inferrule*[right=$\mathtt{DK\_{\to}I{\To}}$]
{\Gamma, \al, \bt, x:\al \vdash e \Lto \bt \dashv \Delta, x:\al, \Theta}
{\Gamma \vdash \lam e \To \al \to \bt \dashv \Delta}
\qquad
\inferrule*[right=$\mathtt{DK\_\forall App}$]
{\Gamma, \al \vdash [\al/a] \appInf{A}{e}{C} \dashv \Delta}
{\Gamma \vdash \appInf{\all A}{e}{C} \dashv \Delta}
$$
In contrast our algorithm collects the existential variables right after the next judgment chain:
\[\Gm \Vdash \lambda x. e \To_a \jg \rrule{25}
\Gm,\al,\bt \Vdash [\al\to\bt/a]\jg, x:\al \Vdash e\Lto \bt\]
\[\Gm \Vdash \appInfAlg{\forall a. A}{e} \rrule{27} \Gm,\al \Vdash \appInfAlg{[\al/a]A}{e}\]
But it seems impossible to achieve a similar outcome by only modifying the two rules in DK's system.
Taking $\mathtt{DK\_{\to}I{\To}}$ as an example,
the declaration or solution for $\al$ and $\bt$ may be referred to by following judgments.
Therefore leaving $\al$ and $\bt$ in the output context is the only choice,
when the next judgments are still inaccessible.

To fix the problem, one possible modification is on the algorithmic subsumption rule,
as garbage collection for a checking judgment is always safe:
$$
\inferrule*[right=$\mathtt{DK\_Sub}$]
{\Gamma, \blacktriangleright_{\al} \vdash e \To A \dashv \Theta \\
\Theta \vdash [\Theta]A \le [\Theta]B \dashv \Delta, \blacktriangleright_{\al}, \Delta'}
{\Gamma \vdash e \Lto B \dashv \Delta}
$$
Here we employ the markers in a way different from what they were originally intended for.
We create a dummy fresh $\al$ and push a marker into the inference judgement.
When the subtyping judgement is used we look for the marker and drop everything afterwards.
We pick this rule because this is the only one where a checking judgment calls an inference judgment.
And that's the point where our algorithm recycles variables---right after a judgment chain is satisfied.
After applying this patch, to our best knowledge,
DK's algorithm behaves equivalently to our algorithm in terms of variable scoping.
However this is exploiting markers in a way that they were not intended for and seems ad-hoc.

@@

Since the judgment form of our algorithmic worklist contains a collection of judgments,
elaboration expressions are also generated as a list of equal length to
the number of judgments (\emph{not judgment chains}) in the worklist.
Within expectation, subtyping judgments translate to coercions (denoted by $f$),
all three other types of judgments translate to terms in System F (denoted by $t$).
\bruno{Why distinguish coercions and terms? Coercions are just terms, right?}

>>>>

One may blame the inference rule for lambda functions for not collecting garbage correctly.
Here we also show another rule for application inference with a similar problem:
$$
\inferrule*[right=$\mathtt{DK\_{\to}I{\To}}$]
{\Gamma, \al, \bt, x:\al \vdash e \Lto \bt \dashv \Delta, x:\al, \Theta}
{\Gamma \vdash \lam e \To \al \to \bt \dashv \Delta}
\qquad
\inferrule*[right=$\mathtt{DK\_\forall App}$]
{\Gamma, \al \vdash [\al/a] \appInf{A}{e}{C} \dashv \Delta}
{\Gamma \vdash \appInf{\all A}{e}{C} \dashv \Delta}
$$
In contrast Rule 25 of our algorithm collects the existential variables
right after the second judgment chain,
and Rule 27 collects one exsitential variable similarly:
\[\Gm \Vdash \lambda x. e \To_a \jg \rrule{25}
\Gm,\al,\bt \Vdash [\al\to\bt/a]\jg, x:\al \Vdash e\Lto \bt\]
\[\Gm \Vdash \appInfAlg{\forall a. A}{e} \rrule{27} \Gm,\al \Vdash \appInfAlg{[\al/a]A}{e}\]
But it seems impossible to achieve a similar outcome by only modifying the two rules in DK's system.
Taking $\mathtt{DK\_{\to}I{\To}}$ as an example,
the declaration or solution for $\al$ and $\bt$ may be referred to by following judgments.
Therefore leaving $\al$ and $\bt$ in the output context is the only choice,
when the next judgments are still inaccessible.

To fix the problem, one possible modification is on the algorithmic subsumption rule,
as garbage collection for a checking judgment is always safe:
$$
\inferrule*[right=$\mathtt{DK\_Sub}$]
{\Gamma, \blacktriangleright_{\al} \vdash e \To A \dashv \Theta \\
\Theta \vdash [\Theta]A \le [\Theta]B \dashv \Delta, \blacktriangleright_{\al}, \Delta'}
{\Gamma \vdash e \Lto B \dashv \Delta}
$$
Here we employ the markers in a way different from what they were originally intended for.
We create a dummy fresh $\al$ and push a marker into the inference judgement.
After the subtyping judgement is processed we look for the marker and drop everything afterwards.
We pick this rule because this is the only one where a checking judgment calls an inference judgment.
And that's the point where our algorithm recycles variables---right after a judgment chain is satisfied.
After applying this patch, to our best knowledge,
DK's algorithm behaves equivalently to our algorithm in terms of variable scoping.
However this is exploiting markers in a way that they were not intended for and seems ad-hoc.

@@

Since the judgment form of our algorithmic worklist contains a collection of judgments,
elaboration expressions are also generated as a list of equal length to
the number of judgments (\emph{not judgment chains}) in the worklist.
Within expectation, subtyping judgments translate to coercions (denoted by $f$),
all three other types of judgments translate to terms in System F (denoted by $t$).
\bruno{Why distinguish coercions and terms? Coercions are just terms, right?}
\jimmy{please help on the question @Tom.}
